Simulation Tables

### 2019-07-02 manuscript/manuscript.tex
algorithm

Mixtures of multivariate non-symmetric distributions such as the multivariate skew-normal (MSN) distribution allow for the nuances of the marginal density to be captured with a more parsimonious set of mixture components. Mixtures of MSN distributions have been dealt with previously in a Bayesian context (Fr\"{u}hwirth-Schnatter \& Pyne, 2010 and others), however in these models, focus lies primary on marginal density estimation and inference on the mixture components (i.e. clusters) is often not discussed. More recently, the mixtures of skew-$t$ factor analysis (MSTFA) model has been proposed for settings in which cluster-specific inference is of primary interest (Lin \textit{et al.} 2018). However, an important feature not included in the MSTFA is the ability to explain individual-level cluster membership as a function of covariates of interest. Additionally, parameter estimation proposed by Lin et al. for the MSTFA relies on a prohibitively complex EM algorithm and does not enjoy the inferential benefits of a Bayesian approach, namely the ability to incorporate prior information into a model and make posterior probability statements. Our proposed model improves on these previous works by estimating parameters in a Bayesian framework as well as including the ability to fit a multinomial logit regression to cluster membership probabilities using a novel application of data augmentation with the P\'olya-Gamma distribution.

@@

\subsection{Statistical Challenges}
\subsubsection{Skewness of Bayley score residuals}

\vspace{1mm}
\begin{figure}[h]
	\caption{The distribution of residuals in repeates measures regression model of Bayley composite scores adjusted for race and sex.}
	\centering
	\includegraphics[scale = 0.14]{bayley_resids_plot}
\end{figure}

\subsubsection{Attrition and Intermittent Missingness}

@@

Lastly, we update $z_1,...,z_n$ by first computing $\boldsymbol\pi_i = (\pi_{1i},...,\pi_{Ki})$ as
$$\pi_{ik} = \frac{\text{e}^{\eta_{ik}}}{1 + \sum_{h=1}^{K-1} \text{e}^{\eta_{ih}}},$$ 
where $\eta_{ik} = \mathbf{w}_i\boldsymbol\delta_k$ for $i = 1,...n$ and $k = 1,...,K$. We also compute according to the multivariate Normal density of $\mathbf{y}_i|t_i$ the probability $P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)$, where $\boldsymbol\zeta_{ki} = \mathbf{x}_i^*\boldsymbol\beta^*_k$. We use these quantities to compute $\boldsymbol\upsilon_{i} = (\upsilon_{1i},...,\upsilon_{Ki})$, where  
$$\upsilon_{ik} = P(z_i = k|\mathbf{y}_i,\boldsymbol\zeta^*_i,\boldsymbol\Sigma_k) = \frac{\pi_{ik}P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)}{\sum_{h = 1}^{K} \pi_{ih}P(\mathbf{y}_i|\boldsymbol\zeta^*_{hi},\boldsymbol\Sigma_h)}$$.
The cluster labels $z_i$ are updated from a $\text{Multinomial}(1,\boldsymbol\varrho_i)$ density for $i = 1,...n$, thus concluding the Gibbs sampler.

@@

A common challenge for Bayesian mixtures models is the so-called ``label switching" problem, in which draws of cluster-specific parameters may be assigned to the wrong cluster at some point during the MCMC simulation, rendering summaries of class-specific parameters incoherent. After conducting simulation studies under a wide variety of realistic parameter settings, as detailed in Section 4, we found little evidence of label switching in posterior draws of $\mathbf{z}$. When label switching was observed, we implemented \textit{post hoc} relabling algorithms as described in Panagiotis 2016 and implemented in the \texttt{label.switching} package in \texttt{R}.

@@

\bibitem{ } Lee SX, Mclachlan GJ. On mixtures of skew normal and skew $t$-distributions. \textit{Advances in Data Analysis and Classification}. 2013 Sep 1;7(3):241-66.

@@

\bibitem{ } Panagiotis Papastamoulis (2016). label.switching: An R Package for Dealing with the Label Switching Problem in MCMC Outputs. \textit{Journal of Statistical Software}, 69(1), 1-24. doi:10.18637/jss.v069.c01

>>>>

Mixtures of multivariate non-symmetric distributions such as the multivariate skew-normal (MSN) distribution allow for the nuances of the marginal density to be captured with a more parsimonious set of mixture components. Mixtures of MSN distributions have been dealt with previously in a Bayesian context (Fr\"{u}hwirth-Schnatter \& Pyne, 2010 and others), however in these models, focus lies primary on marginal density estimation, and inference on the mixture components (i.e. clusters) is often not discussed. More recently, the mixtures of skew-$t$ factor analysis (MSTFA) model has been proposed for settings in which cluster-specific inference is of primary interest (Lin \textit{et al.} 2018). However, an important feature not included in the MSTFA is the ability to explain individual-level cluster membership as a function of covariates of interest. Additionally, parameter estimation proposed by Lin et al. for the MSTFA relies on a prohibitively complex EM algorithm and does not enjoy the inferential benefits of a Bayesian approach, namely the ability to incorporate prior information into a model and make posterior probability statements. Our proposed model improves on these previous works by estimating parameters in a Bayesian framework as well as including the ability to fit a multinomial logit regression to cluster membership probabilities using a novel application of data augmentation with the P\'olya-Gamma distribution.

@@

\subsection{Statistical Challenges}

The analysis of infant development trends in the Nurture data presents a number of statistical challenges that motivate our proposed model. First, as depicted in Figure \ref{fig:skew_resids}, the residuals from repeated measures models of Bayley composite scores exhibit skewness even after adjusting for covariates such as race, sex, and birthweight. This suggests that the assumption of conditional normality made by standard repeated measures models is violated, and a distinguishing feature of the data, namely skewness, is not being accounted for.

\vspace{1mm}
\begin{figure}[h]
	\label{fig:skew_resids}
	\caption{The distribution of residuals in repeates measures regression model of Bayley composite scores adjusted for race and sex.}
	\centering
	\includegraphics[scale = 0.14]{bayley_resids_plot}
\end{figure}

The Nurture data also feature intermittent missingness in Bayley composite scores throughout the study period. Of the total cohort (N = 666), 429 (64.4 \%) observations were available at three months, 435 (65.3 \%) observations were available at six months, 418 (62.8 \%) observations were available at nine months, and 437 (65.6 \%) observations were available at twelve months.

@@

Lastly, we update $z_1,...,z_n$ by first computing $\boldsymbol\pi_i = (\pi_{1i},...,\pi_{Ki})$ as
$$\pi_{ik} = \frac{\text{e}^{\eta_{ik}}}{1 + \sum_{h=1}^{K-1} \text{e}^{\eta_{ih}}},$$ 
where $\eta_{ik} = \mathbf{w}_i\boldsymbol\delta_k$ for $i = 1,...n$ and $k = 1,...,K$. We also compute according to the multivariate Normal density of $\mathbf{y}_i|t_i$ the probability $P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)$, where $\boldsymbol\zeta_{ki} = \mathbf{x}_i^*\boldsymbol\beta^*_k$. We use these quantities to compute $\boldsymbol\upsilon_{i} = (\upsilon_{1i},...,\upsilon_{Ki})$, where  
$$\upsilon_{ik} = P(z_i = k|\mathbf{y}_i,\boldsymbol\zeta^*_i,\boldsymbol\Sigma_k) = \frac{\pi_{ik}P(\mathbf{y}_i|\boldsymbol\zeta^*_{ki},\boldsymbol\Sigma_k)}{\sum_{h = 1}^{K} \pi_{ih}P(\mathbf{y}_i|\boldsymbol\zeta^*_{hi},\boldsymbol\Sigma_h)}$$.
The cluster labels $z_i$ are updated from a $\text{Multinomial}(1,\boldsymbol\upsilon_i)$ density for $i = 1,...n$, thus concluding the Gibbs sampler.

@@

A common challenge for Bayesian mixtures models is the so-called ``label switching" problem, in which draws of cluster-specific parameters may be assigned to the wrong cluster at some point during the MCMC simulation, rendering summaries of class-specific parameters incoherent. After conducting simulation studies under a wide variety of realistic parameter settings, as detailed in Section 4, we found little evidence of label switching in posterior draws of $\mathbf{z}$. When label switching was observed, we implemented \textit{post hoc} relabling algorithms as described in Papastamoulis 2016 and implemented in the \texttt{label.switching} package in \texttt{R}.

@@

\bibitem{ } Lee SX, Mclachlan GJ. On mixtures of skew normal and skew-$t$ distributions. \textit{Advances in Data Analysis and Classification}. 2013 Sep 1;7(3):241-66.

@@

\bibitem{ } Papastamoulis, P (2016). label.switching: An R Package for Dealing with the Label Switching Problem in MCMC Outputs. \textit{Journal of Statistical Software}, 69(1), 1-24. doi:10.18637/jss.v069.c01

### 2019-07-02 mcmc_draws_2019-07-02/tables.Rmd
algorithm

```{r}
setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-01")
load("BETA")
load("beta_true")
load("DELTA")
load("delta_true")
load("PI")
load("PSI")
load("psi_true")
load("SIGMA")
load("sig2_true")
load("pi_true")
load("Z")
```

>>>>

```{r}
setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-02")
load("BETA")
load("beta_true")
load("DELTA")
load("delta_true")
load("PI")
load("PSI")
load("psi_true")
load("SIGMA")
load("sig2_true")
load("pi_true")
load("Z")
```
